\documentclass{article}

\usepackage{amsmath}

\begin{document}

\title{Fall 2018 14.01 Problem Set 2}
\maketitle

\section*{Problem 1: True or False (24 points)}
For each of the following statements, indicate if they are True or False. Justify your answer. 

\begin{enumerate}
    \item (4 points) Suppose potatoes are a Giﬀen good. When the price of potatoes increases, both the substitution and the income eﬀects cause the demand for potatoes to increase.
    \item (4 points) John consumes only sodas (x) and pizza (y). Suppose that his preferences can be represented by $U = \min \{x, y\}$, so sodas and pizza are perfect complements: John wants to drink one soda for every pizza he eats. If the price of sodas increases, John’s demand for both sodas and pizzas will decrease, and all the change in demand will be due to the substitution eﬀect.
    \item (4 points) Consider a situation with only two goods, x and y. It is possible for one of the goods to be inferior, but they can’t both be inferior at the same time.
    \item (4 points) A production function uses capital and labor as inputs. If both capital and labor have diminishing marginal returns, then the production function can’t display increasing returns to scale.
    \item (4 points) Returns to scale for a ﬁrm are the same at all production levels.
    \item (4 points) Thomas Malthus’ prediction about mass starvation was wrong because the diminishing marginal product of labor has been oﬀset by the increase in agricultural productivity.
\end{enumerate}

\section*{Problem 2 (35 points)}
Anne consumes only books (x) and video games (y). Her preferences can be represented by the following utility function: $U = x^2y$. The price of books is $px$, the price of video games is $py$, and Anne has an income of $m$ dollars. 

\begin{enumerate}
    \item (5 points) Write down Anne’s budget constraint, calculate the Marginal Rate of Substitution (at an arbitrary bundle ($x$, $y$)) and compute her demand for books and video games (as a function of $px$, $py$ and $m$).
    \item (5 points) Compute the price elasticity of the demand for books. Compute the cross-price elasticity of the demand for books with respect to the price of video games. (The cross-price elasticity measures the change of the quantity demanded for a good in response to a change in the price of another good, holding everything else constant. It is measured as the percentage change in quantity demanded for the ﬁrst good divided by the percentage change in price of the second good.)
    
    % Further items...
\end{enumerate}

% Further problems...

\end{document}