\documentclass{article}
\usepackage{amsmath} 

\begin{document}
\title{Fall 2018 14.01 Problem Set 2 - Solutions}
\maketitle

\section{Problem 1: True or False (24 points)}
For each of the following statements, indicate if they are True or False. Justify your answer.

\begin{enumerate}
\item(4 points) Suppose potatoes are a Giffen good. When the price of potatoes increases, both the substitution and the income effects cause the demand for potatoes to increase.

Solution: False, the substitution effect always leads to a decrease in demand, but in the case of a Giffen good the income effect is sufficiently strong to lead to an increase in demand.

\item(4 points) John consumes only sodas (x) and pizza (y) . Suppose that his preferences can be represented by $U = \min \{x, y\} $, so sodas and pizza are perfect complements: John wants to drink one soda for every pizza he eats. If the price of sodas increases, John's demand for both sodas and pizzas will decrease, and all the change in demand will be due to the substitution effect.
Solution: False, all the change in demand is due to income effect.
\item(4 points) Consider a situation with only two goods, x and y. It is possible for one of the goods to be inferior, but they can't both be inferior at the same time.
Solution: True, it is not possible for both goods to be inferior at the same time. If that were the case, when income increases the consumer would spend a lower total amount of money than before, which contradicts non-satiation. Note: extra credit to be given if non-satiation is explicitly mentioned as a necessary assumption for this to hold.
\end{enumerate}
\end{document}