\documentclass{article}
\usepackage{amsmath}

\title{Fall 2018 14.01 Problem Set 5}
\date{}
\begin{document}
\maketitle

\noindent\textbf{Problem 1: True or False (24 points).} \\
Determine whether the following statements are True or False. Explain your answer. 

\begin{enumerate}
\item (4 points) A government sets a price ceiling for widgets that is below the equilibrium price. This intervention will always decrease the producer surplus, increase consumer surplus and decrease total surplus. 

\item (4 points) A non-discriminating monopoly will optimally choose a price and quantity where the market demand curve is inelastic (i.e. elasticity is lower than one in absolute value).

\item (4 points) If a monopolist charging a uniform price begins to practice perfect price discrimination, some consumers may become strictly better off.

\item (4 points) Transferring a monopolist’s profit to consumers eliminates the inefficiency associated with monopoly.

\item (4 points) Martin knows that if he sells 500 widgets, his revenue will be \$1000 and that if he sells 800 widgets, his revenue will be \$1500. The market for widgets is perfectly competitive.

\item (4 points) Suppose that a monopolist initially charging a uniform price, now moves to charging a different price to young and old consumers, but the total quantity sold by the monopolist doesn’t change. Claim: total welfare must be lower after the monopolist starts discriminating. 
\end{enumerate}

\begin{flushleft}
... [truncated for brevity] ...
\end{flushleft}

\end{document}