\documentclass[12pt]{article}
\title{Fall 2018 14.01 Problem Set 1}
\date{}
\begin{document}
\maketitle
\noindent
\section*{1 Positive vs. Normative Statements (16 points)}
Identify whether each of the following statements is positive or normative. Briefly\\ 
justify your answer.
\begin{enumerate}
\item (4 points) The government has a duty to provide basic healthcare and education to every citizen.
\item (4 points) The cost of health insurance is too high.
\item (4 points) The median earnings of a full-time worker with a college degree are almost twice as high as those of a high-school graduate with no college education.
\item (4 points) The current unemployment rate is 3.9\%, the lowest it has been since December 2000.
\end{enumerate}

\section*{2 True or False (20 points)}
For each of the following statements, indicate if they are True or False. Justify your answer.
\begin{enumerate}
\item Bill is a football coach...
    \begin{enumerate}
    \item (4 points) Bill’s preferences are complete.
    \item (4 points) Bill’s preferences are transitive.
    \end{enumerate}
\item (4 points) Ann and Bob are utility maximizing consumers...
\item (4 points) John’s utility function for food (f) and clothes (c) is given by $U (f, c) = \alpha (f^\alpha + c^\alpha)^{1/\alpha}$, with $\alpha > 0$. 
\item (4 points) Ava has preferences over two goods...
\end{enumerate}

\section*{3 Demand for Video Games (16 points)}
We have the following weekly demand data...

\section*{4 Indifference Curves (18 points)}
In each of the following examples...

\section*{5 Utility Maximization (30 points)}
Chloe consumes only books (x) and video games (y) . Her preferences...

\noindent
MIT OpenCourseWare https://ocw.mit.edu\\
14.01 Principles of Microeconomics\\
Fall 2018\\
For information about citing these materials or our Terms of Use, visit: https://ocw.mit.edu/terms.

\end{document}